\begin{table}[t]
	\centering
	\begin{smalltrees}
		\begin{tabular}{ccccccc}
    		\cline{1-3}\cline{5-7}
    		$i$ & tree $t_i$ & elementary weight & & $i$ & tree $t_i$ & elementary weight \\
    		\cline{1-3}\cline{5-7} \\[-10pt]
    		0 & $\emptyset$ \hspace{15pt}  & 1 & & 9 & \hspace{15pt} \tree{9} & $\bm{b}^T\bm{c}^4$\\
    		1 & \hspace{15pt}  \tree{1} &$\bm{b}^T\bm{e}$ & & 10 & \tree{10} \hspace{15pt} & $\bm{b}^TC^2A\bm{c}$ \\
    		2 & \tree{2} \hspace{15pt}  &$\bm{b}^T\bm{c}$ & & 11 & \hspace{15pt} \tree{11} & $\bm{b}^TCA\bm{c}^2$ \\
    		3 & \hspace{15pt}  \tree{3} & $\bm{b}^T\bm{c}^2$ & & 12 & \tree{12} \hspace{15pt} & $\bm{b}^TCA^2\bm{c}$ \\
    		4 & \tree{4} \hspace{15pt}  & $\bm{b}^TA\bm{c}$ & & 13 & \hspace{15pt} \tree{13} & $\bm{b}^T(A\bm{c})^2$ \\
    		5 & \hspace{15pt}  \tree{5} & $\bm{b}^T\bm{c}^3$ & & 14 & \tree{14} \hspace{15pt} & $\bm{b}^TA\bm{c}^3$ \\
    		6 & \tree{6} \hspace{15pt}  & $\bm{b}^TCA\bm{c}$ & & 15 & \hspace{15pt} \tree{15} & $\bm{b}^TACA\bm{c}$ \\
    		7 & \hspace{15pt}  \tree{7} & $\bm{b}^TA\bm{c}^2$ & & 16 & \tree{16} \hspace{15pt} & $\bm{b}^TA^2\bm{c}^2$ \\
    		8 & \tree{8} \hspace{15pt}  & $\bm{b}^TA^2\bm{c}$ & &  17 & \hspace{15pt} \tree{17} & $\bm{b}^TA^3\bm{c}$ \\
  		\end{tabular}
  \end{smalltrees}
  \caption{Elementary weights of trees up to fifth order for a 
  		Runge--Kutta method $(A,\bm{b},\bm{c})$. 
  		Here $C$ is a diagonal matrix with components 
  		$c_{i} = \sum_j^s a_{ij}$ and exponents of vectors 
  		represent component exponentiation.}
  \label{tab:elementary_weights}
\end{table}

\section{Effective order Runge--Kutta methods}\label{sec:Algebraic_RK}
The effective order of a Runge--Kutta method is defined in an abstract 
algebraic context introduced by Butcher \cite{Butcher1969} and developed 
further in \cite{Butcher1972, Hairer1974, Butcher1996, Butcher1998} and 
others.
In this section we follow \cite{Butcher2008_book}, reviewing the 
fundamental concepts of this representation which are then used to 
define effective order methods and their order conditions.

\subsection{The Runge--Kutta group}\label{sec:RK_group}
%The Runge--Kutta methods can be viewed as elements of an algebraic 
%group in which the product is related to the composition of two methods.
%Butcher \cite{Butcher1972} proposed the group $G$ of all real-valued 
%maps on the set of rooted trees.
%\davidtodo{Did he propose the group (whatever that means),
%or did he show that it is isomorphic to the Runge--Kutta group?}\yiannistodo{He did both, de defined $G$ and proved is isomorphic to the 
%composition group of RK methods. We had a long discussion with Colin on 
%whether mentioning the composition group and decided not.}
%Each function $\alpha \in G$ corresponds to an \emph{equivalence class} 
%of Runge--Kutta methods and maps trees to specific algebraic expressions 
%in the coefficients of a Runge--Kutta method, known as 
%\emph{elementary weights}.
%%The elementary weight of a tree $t$ is an algebraic expression 
%%in the coefficients of a Runge--Kutta method.
%Two Runge--Kutta methods belong to the same equivalence class 
%(and hence called \emph{equivalent}) if they have the same elementary weight values.

Runge--Kutta methods can be viewed as elements of an algebraic group
in which the product corresponds to the composition of two methods.
%corresponds to successive application of two methods.
Let $G$ be the group of all real-valued maps on the set
of rooted trees.
Each function $\alpha \in G$ corresponds
to an \emph{equivalence class} of Runge--Kutta methods and maps trees to specific
algebraic expressions in the coefficients
of a Runge--Kutta method, known as \emph{elementary weights}.
%The elementary weight of a tree $t$ is an algebraic expression in the coefficients
%of a Runge--Kutta method.
Two Runge--Kutta methods belong to the same equivalence class 
if they have the same elementary weight values.

For every function $\alpha \in G$ we write the values of the
elementary weights as $\alpha_{i} = \alpha(t_{i})$ for trees $t_{i}$
indexed by integer $i$.
The vector of these values $\alpha_i$, $i = 0, 1, 2,\ldots$ is referred to 
as the B-series of the corresponding Runge--Kutta method and is 
related to Taylor expansions of the numerical solution given by 
the Runge--Kutta method \cite{Hairer1974, Butcher2008_book}. 
By convention $\alpha(t_{0}) = 1$, where $t_{0}$ denotes the empty tree. 
Table~\ref{tab:elementary_weights} lists these expressions for trees of 
up to degree five; a general recursive formula can be found in 
\cite[Definition 312A]{Butcher2008_book}. 

% \subsubsection*{minimalist}
% Given $\alpha, \beta \in G$, the product can be defined on each
% tree by performing a certain partitioning of the tree and computing
% over the resulting forest \cite{Butcher2008_book}.  The product of
% $\alpha\beta$ corresponds to the application of one Runge--Kutta
% method followed by another.  This multiplicative binary operation
% makes $G$ a group.

Let $\alpha, \beta \in G$ correspond to Runge--Kutta methods A and B
respectively.
%Let $\alpha, \beta \in G$ correspond to two Runge--Kutta
%equivalence classes and assume $A$ and $B$ to be representative
%methods of each group.
A multiplicative group operation $\alpha\beta$ can be defined
by partitioning the input tree and computing over the resulting 
forest \cite{Butcher2008_book}.
This product $\alpha\beta$ is related to the application
of method $A$ followed by method $B$; 
we denote the resulting method by $BA$.\footnote{We write
	$BA$ to mean the application of method $A$
	followed by the application of method $B$ 
	(following matrix and operator ordering convention) 
	but when referring to products of elements of the
	Runge--Kutta group we use the reverse ordering ($\alpha\beta$) 
	to match the convention in \cite{Butcher2008_book}.}
The product is defined by
\begin{equation}\label{eq:Group_operation}
	(\alpha\beta)(t) = \sum_{w \lhd t} \biggl(\prod_{v \in t \setminus w} \alpha(v)\beta(w)\biggr),
\end{equation}
where $w \lhd t$ indicates a subtree of $t$ which includes the
root of $t$ and $w \setminus t$ indicates the forest induced
by removing $w$ from $t$ \cite{Butcher2008_book}.
Multiplicity in choosing $w$ must also be accounted for.
The following example shows how to compute this product for
one particular tree.

%\subsubsection*{longer version}
% The group multiplicative operation $\alpha\beta$ corresponds to the
% application of one Runge--Kutta method followed by another.
% It is defined as follows.
% \begin{definition}\label{def:Group_operation}
% 	Let $\alpha$ and $\beta$ be two members of $G$, mapping trees to real numbers. Then for every tree $t$, we define their product by
% 	\begin{equation}\label{eq:Group_operation}
% 		(\beta\alpha)(t) = \sum_{w \lhd t} \prod_{v \in t \setminus w} \alpha(v)\beta(w),
% 	\end{equation}
%         where $w \lhd t$ indicates a subtree of $t$ which includes the
%         root of $t$ and $w \setminus t$ indicates the forest induced
%         by removing $w$ from $t$.
%         Multiplicity in choosing $w$ must also be accounted for.
% \end{definition}
% See \cite{Butcher2008_book} for details.  This product is perhaps best
% understood via an example.

\begin{example}\label{ex:tree_partition}
	Table~\ref{tab:tree_partition} shows the partition of the 
	five-vertex tree $t_{11}$ to all possible rooted subtrees. 
	Based on this partition, we apply \eqref{eq:Group_operation} to 
	find that the product of two functions in $G$ on tree $t_{11}$ is
	given by $(\alpha\beta)(t_{11}) = \alpha_{11} + 
	\alpha_1\alpha_3\beta_1 + (\alpha_1^{3} + \alpha_3)\beta_2 + 
	\alpha_1^{2}\beta_3 + 2\alpha_1^{2}\beta_4 + 2\alpha_1\beta_6 + 
	\alpha_1\beta_7 + \beta_{11}.$
	\begin{table}
		\centering
    		\begin{tabular}{c cc|c|c|c|c|c|c|c|c|c}
        		\multirow{4}{*}{\begin{largetrees}\treecell{$\tree{11}$}{$t_{11}$}\end{largetrees}} & & 
        		$w$ & $\emptyset$ & $\tree{1}$ & $\tree{2}$ & $\tree{2}$ & $\tree{3}$ & $\tree{4}$ 
        		& $\tree{6}$ & $\tree{7}$ & $\tree{11}$ \\[3pt]
	        \cline{3-12}
	        & & $t_{11} \setminus w$ & \rowscell{$\tree{11}$}{} & \rowscell{$\tree{1}$}{$\tree{3}$} 
	        & \rowscell{$\tree{3}$}{ } & \rowscell{$\tree{1} \quad \tree{1}$}{$\tree{1}$} & 
	        \rowscell{$\tree{1} \quad \tree{1}$}{ } & \rowscell{$\tree{1} \quad \tree{1}$}
	        {$(\times2)$} & \rowscell{$\tree{1}$}{$(\times2)$} & \rowscell{$\tree{1}$}{ } & 
	        \rowscell{$\emptyset$}{ }
	     \end{tabular}
	     \vspace{5pt}
	     \caption{Partitions of tree $t_{11}$ to all possible subtrees $w$ 
	     	and the corresponding forests $t_{11} \setminus w$. 
	     	Multiplicity is indicated with $(\times2)$.}
	     \label{tab:tree_partition}
	\end{table}
\end{example}

\subsection{Algebraic interpretation of order}\label{sec:Algebraic_order}
If two Runge--Kutta methods correspond to the same element in $G$, 
then they are essentially the same method (up to reducibility).
%For our purposes, we want a weaker condition: a way to
%discuss equivalence of methods up to a particular order of accuracy.
However, the definition of equivalence of methods is overly restrictive 
for practical purposes. A weaker condition is established if we 
discuss equivalence of methods up to a particular order of accuracy.
\begin{definition}\label{def:Equivalent_methods}
	Two Runge--Kutta methods $A$ and $B$, are equivalent up to order 
	$p$ if their corresponding elements in $G$, $\alpha$ and $\beta$, satisfy
	\begin{align*}
		\alpha(t) = \beta(t), \; \text{for every tree $t$ with $r(t) \leq p$},
	\end{align*}
	where $r(t)$ denotes the order of the tree (number of vertices).
	We denote this equivalence relation by $A \equival{p} B$. 
\end{definition}

In this sense, methods have inverses: the product of $\alpha^{-1}$ and
$\alpha$ must match the identity method up to order $p$.
Classical order follows from comparing a method with the special group
element $E \in G$ which advances the exact solution by one step.
All of this can be made considerably more precise using quotient
groups of $G$ \cite{Butcher2008_book}.
\begin{example}\label{ex:FE_inv_2}
  Consider the forward Euler method
  $\bm{u}^{n+1} = \bm{u}^{n} + \Dt \bm{F}(\bm{u}^{n})$.
  To find an inverse, we seek a method that undoes the work of this method,
  recovering $\bm{u}^{n}$ from $\bm{u}^{n+1}$;
  one approach is to solve for $\bm{u}^{n}$, obtaining the backward Euler method
  with a time-step of $-\Dt$.
  Alternatively, let $\alpha \in G$ correspond to the forward Euler method and by
  \eqref{eq:Group_operation}, we have
  %\cite[Sec~382]{Butcher2008_book}, we have
  $(\alpha\alpha^{-1})(t_1) = \alpha(t_1) + \alpha^{-1}(t_1) = 0$,
  so any $\alpha^{-1}$ with $\alpha^{-1}(t_1) = -1$ will do.
  For example, the forward Euler method with size $-\Delta t$ is also an
  inverse (up to order $1$).
\end{example}
This example demonstrates that inverse methods up to order $p$ are not
unique and inverse methods of explicit methods need not be implicit.

\subsection{Effective order}\label{sec:Effective_order}
Effective order is achieved by using a \emph{starting method} $S$
followed by a \emph{main method} $M$
and then a \emph{stopping method} $S^{-1}$.
%: i.e., the inverse of $S$ which
%annihilates the work of the starting method (up to order $q$).
%The methods $S$ and $S^{-1}$ make the method $M$ effectively of higher order than its classical order.
We denote by $\alpha$ and $\beta$ the elements of $G$ associated with the methods $M$ and $S$, respectively.
%Thus $\beta^{-1}$ corresponds to Runge--Kutta method $S^{-1}$ that annihilates the work of the starting method $S$ (up to order $q$).
The successive use of these three methods results in a method $P = S^{-1}MS$, of which the corresponding element of $G$ is $\beta\alpha\beta^{-1}$.
We want $P$ to have order $q$ whereas $M$ might have lower classical
order $p < q$.
In terms of functions in group $G$ this is leads to the following definition of the effective order of the Runge--Kutta method $M$.
\begin{definition}\cite{Butcher1987_book}\label{def:Effective_order}
  % Let $\alpha$ and $\beta be the corresponding functions in $G$ of a
  % Runge--Kutta methods $M$ and $S$, respectively.
  Suppose $M$ is a Runge--Kutta method with corresponding $\alpha \in G$.
  Then the method $M$ is of effective order $q$ if there exists a method
  $S$ (with corresponding $\beta \in G$) such that
	\begin{equation}\label{eq:Effective_order_1}
		(\beta\alpha\beta^{-1})(t) = E(t), \; \text{for every tree with $r(t) \leq q$,}
	\end{equation}
        where $\beta^{-1}$ is an inverse of $\beta$ up to order $q$.
        Recall that $E$ represents one exact step of the solution.
\end{definition}
The practical benefit of methods of effective order results from the
observation that only $M$ need be used repeatedly, since
\begin{displaymath}
	P^n = (S^{-1}MS)^n = (S^{-1}MS) \cdots (S^{-1}MS) (S^{-1}MS) \equival{q} S^{-1} M^n S.
\end{displaymath}
The starting method is applied at the beginning without advancing the
solution.
Instead, it introduces a perturbation to the solution.
The main method  $M$ is then used for $n$ time steps and finally the
stopping method is used to correct the solution.
In Section~\ref{subsec:starting_stopping}, we propose alternative
starting and stopping procedures which allow the overall procedure to
be SSP.

%\begin{result}
%  Let method $M$ have effective order $q$ with associated methods $S$
%  and $S^{-1}$.  Let $P=S^{-1}MS$ then
%  $$P^n = (S^{-1}MS)^n = (S^{-1}MS) \cdots (S^{-1}MS) (S^{-1}MS)
%        = S^{-1} M^n S.$$
%The starting method is applied at the beginning without advancing the
%solution.
%Instead, it introduces a perturbation on the solution.
%The main method \( M \) is then used for \( n \) time steps and finally the
%stopping method is used to correct the solution.
%%Notably, if the classical order of $M$ is $p < q$, (insert local truncation error bit)...
%\end{result}

\subsubsection{Effective order conditions}\label{sec:effOrderCond}
For the main method $M$ to have effective order $q$ its coefficients must satisfy a set of algebraic conditions corresponding to the rooted trees of order up to $q$.
That is, the Runge--Kutta method $M$ corresponding to the function $\alpha$ must satisfy
\emph{effective order conditions} relative to the order conditions of the
method $S$ corresponding to the function $\beta$.
We rewrite \eqref{eq:Effective_order_1} as
$(\beta\alpha)(t) = (E\beta)(t), \; \text{for all trees with $r(t) \leq q$,}$
and using the product operation \eqref{eq:Group_operation}, we can find expressions for each tree $t$ with $r(t) \leq q$.
For trees up to order five these are tabulated in \cite[Sec~389]{Butcher2008_book}
and in Table~\ref{tab:effective_OCs_on_alpha}.
In general, the effective order conditions allow more degrees of
freedom on methods than the classical order conditions.
Note that these conditions match the classical order conditions up to
second order.
Note also that for the tall trees $t_1, t_4, t_8, t_{17}, \dots$ the
effective order conditions of the main method match the classical
order conditions and that these are precisely the order conditions
for linear problems \cite{Butcher2008_book}.
\begin{table}
	%\vspace*{-5ex}
	%\setlength{\columnseprule}{0.4pt}
	%\setlength{\columnsep}{10pt}
	\begin{multicols}{2}
		\begin{align*}
			% tfrac used here to make smaller fractions
    		\alpha_1  &= 1 \\
     		\alpha_2  &= \tfrac{1}{2} \\
    		\alpha_3  &= \tfrac{1}{3} + 2\beta_2 \\
    		\alpha_4  &= \tfrac{1}{6} \\
    		\alpha_5  &= \tfrac{1}{4} + 3\beta_2 + 3\beta_3 \\
    		\alpha_6  &= \tfrac{1}{8} + \beta_2 + \beta_3 + \beta_4 \\
    		\alpha_7  &= \tfrac{1}{12} +\beta_2 - \beta_3 + 2\beta_4 \\
    		\alpha_8  &= \tfrac{1}{24}
    	\end{align*}
    	\vfill
    	\columnbreak
    	\begin{align*}
    		\alpha_9  &= \tfrac{1}{5} + 4\beta_2 + 6\beta_3 + 4\beta_5 \\
    		\alpha_{10} &= \tfrac{1}{10} + \tfrac{5}{3}\beta_2 - 2\beta_2^{2} + \tfrac{5}{2}\beta_3 + \beta_4 + \beta_5 + 2\beta_6 \\
    		\alpha_{11} &= \tfrac{1}{15} + \tfrac{4}{3}\beta_2 + \tfrac{1}{2}\beta_3 + 2\beta_4 + 2\beta_6 + \beta_7 \\
    		\alpha_{12} &= \tfrac{1}{30} + \tfrac{1}{3}\beta_2 - 2\beta_2^{2} + \tfrac{1}{2}\beta_3 + \tfrac{1}{2}\beta_4 + \beta_6 + \beta_8 \\
    		\alpha_{13} &= \tfrac{1}{20} + \tfrac{2}{3}\beta_2 - \beta_2^{2} + \beta_3 + \beta_4 + 2\beta_6 \\
    		\alpha_{14} &= \tfrac{1}{20} + \beta_2 + 3\beta_4 - \beta_5 + 3\beta_7 \\
    		\alpha_{15} &= \tfrac{1}{40} + \tfrac{1}{3}\beta_2 + \tfrac{3}{2}\beta_4 - \beta_6 + \beta_7 + \beta_8 \\
    		\alpha_{16} &= \tfrac{1}{60} + \tfrac{1}{3}\beta_2 - \tfrac{1}{2}\beta_3 + \beta_4 - \beta_7 + 2\beta_8 \\
    		\alpha_{17} &= \tfrac{1}{120}
    	\end{align*}%
        %\vfill
    \end{multicols}
    \caption{Effective order five conditions of the main and starting methods $M$ and $S$.}
    \label{tab:effective_OCs_on_alpha}
\end{table}
%Finally, we use the abbreviation RK($s$,$q$,$p$) for an $s$-stage Runge-Kutta method of effective order $q$ and classical order $p$.


\subsection{Constructing effective order methods}
The approach we adopt is to consider the $\beta_{i}$ as free
parameters when determining the $\alpha_i$.
The relationship in Table~\ref{tab:effective_OCs_on_alpha} between the
$\alpha_i$ and $\beta_i$ is mostly linear (although there are a few
$\beta_2^2$ terms).
It is thus straightforward to isolate the equations for $\alpha_i$,
determining the $\beta_i$ as linear combination of the $\alpha_i$, and
separate the effective order conditions into conditions on main method
$M$ and starting method $S$.
This provides maximal degrees of freedom and minimizes the number of
constraints when constructing the method $M$.
Then when all functions on $\alpha$ are found, the relative order
conditions on $\beta$ can be obtained.

The resulting effective order conditions for main method $M$ are given
in Table~\ref{tab:effective_OCs} (up to effective order five).
The order conditions for the starting method $S$ are also given.
We can also find the order conditions of
%$\beta^{-1}$
$S^{-1}$
in terms of the
$\beta_i$ (see \cite[Table~386(III)]{Butcher2008_book}).

Note that we follow the convention of setting $\beta_1=0$
\cite{Butcher2008_book}, i.e., the starting and stopping methods
perturb the solution but do not advance the solution in time
(Tables~\ref{tab:effective_OCs_on_alpha} and~\ref{tab:effective_OCs} both incorporate
this assumption).

\begin{table}[htb]
	\centering
    \begin{tabular}{M{2mm}|M{2mm}|M{68mm}|M{67mm}}
    		\hline
        q & p & Order conditions for the main method $M$ & Order conditions for the starting method $S$ \nline
        \hline
        \multirow{1}{*}{3} & \multirow{1}{*}{2} & {\small $\alpha_1 = 1$, $\alpha_2 = \frac{1}{2}$, $\alpha_4 = \frac{1}{6}$.} & {\small $\beta_1 = 0$, $\beta_2 = - \frac{1}{6} + \frac{1}{2}\alpha_3$.}\nline
        \hline
        \multirow{3}{*}{4} & \multirow{3}{*}{2} & {\small $\alpha_1 = 1$, $\alpha_2 = \frac{1}{2}$, $\alpha_4 = \frac{1}{6}$,} & {\small $\beta_1 = 0$, $\beta_2 = - \frac{1}{6} + \frac{1}{2}\alpha_3$,}\nline
        & & {\small $\frac{1}{4} - \alpha_3 + \alpha_5 - 2\alpha_6 + \alpha_7 = 0$, $\alpha_8 = \frac{1}{24}$.} & {\small $\beta_3 = \frac{1}{12} - \frac{1}{2}\alpha_3 + \frac{1}{3}\alpha_5$, $\beta_4 = - \frac{1}{24} - \frac{1}{3}\alpha_5 + \alpha_6$.} \nline
        \hline
        \multirow{3}{*}{4} & \multirow{3}{*}{3} & {\small $\alpha_1 = 1$, $\alpha_2 = \frac{1}{2}$, $\alpha_3 = \frac{1}{3}$, $\alpha_4 = \frac{1}{6}$,} & {\small $\beta_1 = 0$, $\beta_2 = 0$, $\beta_3 = - \frac{1}{12}  + \frac{1}{3}\alpha_5$,} \nline
        & & {\small $\frac{1}{12} - \alpha_5 + 2\alpha_6 - \alpha_7 = 0$, $\alpha_8 = \frac{1}{24}$.} & {\small $\beta_4 = - \frac{1}{24} - \frac{1}{3}\alpha_5 + \alpha_6$.} \nline
        \hline
        \multirow{8}{*}{5} & \multirow{8}{*}{2} & {\small $\alpha_1 = 1$, $\alpha_2 = \frac{1}{2}$, $\alpha_4 = \frac{1}{6}$, $\alpha_8 = \frac{1}{24}$, $\alpha_{17} = \frac{1}{120}$,} & {\small $\beta_1 = 0$, $\beta_2 = - \frac{1}{6} + \frac{1}{2}\alpha_3$,} \nline
        & & {\small $\frac{1}{4} - \alpha_3 + \alpha_5 - 2\alpha_6 + \alpha_7 = 0$,} & {\small $\beta_3 = \frac{1}{12} - \frac{1}{2}\alpha_3 + \frac{1}{3}\alpha_5$, $\beta_4 = -\frac{1}{24} - \frac{1}{3}\alpha_5 + \alpha_6$} \nline
        & & {\small $\frac{1}{4}\alpha_9-\alpha_{10}+\alpha_{13}=\beta_2^{2}$, \: $\beta_2 = - \frac{1}{6} + \frac{1}{2}\alpha_3$,} & {\small $\beta_5 = -\frac{1}{120} + \frac{1}{4}\alpha_3 - \frac{1}{2}\alpha_5 + \frac{1}{4}\alpha_9$,} \nline
        & & {\small $\frac{3}{10} - \frac{3}{2}\alpha_3 + \alpha_5 + \frac{1}{2}\alpha_9 - 3\alpha_{10} + 3\alpha_{11} - \alpha_{14} = 6\beta_2^{2}$,} & {\small $\beta_6 = \frac{7}{720} + \beta_2^{2} + \frac{1}{12}\alpha_3 - \frac{1}{2}\alpha_6 - \frac{1}{8}\alpha_9 + \frac{1}{2}\alpha_{10}$,} \nline
        & & {\small $\frac{1}{15} - \frac{1}{2}\alpha_3 + \alpha_6 + \frac{1}{2}\alpha_9 - 2\alpha_{10} + \alpha_{11} + \alpha_{12} - \alpha_{15} = 2\beta_2^{2}$,} & {\small $\beta_7 = \frac{8}{45} - 2\beta_2^{2} - \frac{7}{12}\alpha_3 + \frac{1}{2}\alpha_5 - \alpha_6 + \frac{1}{4}\alpha_9 - \alpha_{10} + \alpha_{11}$,} \nline
        & & {\small $\frac{19}{60} - \alpha_3 + \alpha_5 - 2\alpha_6 + \alpha_{11} - 2\alpha_{12} + \alpha_{16} = 4\beta_2^{2}$.} & {\small $\beta_8 = -\frac{1}{120} + \beta_2^{2} + \frac{1}{8}\alpha_9 - \frac{1}{2}\alpha_{10} + \alpha_{12}$.} \nline
        \hline
        \multirow{7}{*}{5} & \multirow{7}{*}{3} & {\small $\alpha_1 = 1$, $\alpha_2 = \frac{1}{2}$, $\alpha_3 = \frac{1}{3}$, $\alpha_4 = \frac{1}{6}$, $\alpha_8 = \frac{1}{24}$,} & {\small $\beta_1 = 0$, $\beta_2 = 0$, $\beta_3 = -\frac{1}{12} + \frac{1}{3}\alpha_5$} \nline
        & & {\small $\alpha_{17} = \frac{1}{120}$, $\frac{1}{12} - \alpha_5 + 2\alpha_6 - \alpha_7 = 0$,} & {\small $\beta_4 = -\frac{1}{24} - \frac{1}{3}\alpha_5 + \alpha_6$,} \nline
        & & {\small $\frac{1}{4}\alpha_9 - \alpha_{10} + \alpha_{13} = 0$,} & {\small $\beta_5 = \frac{3}{40} - \frac{1}{2}\alpha_5 + \frac{1}{4}\alpha_9$,} \nline
        & & {\small $\frac{1}{5} - \alpha_5 - \frac{1}{2}\alpha_9 + 3\alpha_{10} - 3\alpha_{11} + \alpha_{14} = 0$,} & {\small $\beta_6 = \frac{3}{80} - \frac{1}{2}\alpha_6 - \frac{1}{8}\alpha_9 + \frac{1}{2}\alpha_{10}$,} \nline
        & & {\small $\frac{1}{10} - \alpha_6 - \frac{1}{2}\alpha_9 + 2\alpha_{10} - \alpha_{11} - \alpha_{12} + \alpha_{15} = 0$,} & {\small $\beta_7 = -\frac{1}{60} + \frac{1}{2}\alpha_5 - \alpha_6 + \frac{1}{4}\alpha_9 - \alpha_{10} + \alpha_{11}$,} \nline
        & & {\small $\frac{1}{60} - \alpha_5 + 2\alpha_6 - \alpha_{11} + 2\alpha_{12} - \alpha_{16} = 0$.} & {\small $\beta_8 = -\frac{1}{120} + \frac{1}{8}\alpha_9 - \frac{1}{2}\alpha_{10} + \alpha_{12}$.} \nline
        \hline
        \multirow{7}{*}{5} & \multirow{7}{*}{4} & {\small $\alpha_1 = 1$, $\alpha_2 = \frac{1}{2}$, $\alpha_3 = \frac{1}{3}$, $\alpha_4 = \frac{1}{6}$, $\alpha_5 = \frac{1}{4}$,} & {\small $\beta_1 = 0$, $\beta_2 = 0$,} \nline
        & & {\small $\alpha_6 = \frac{1}{8}$, $\alpha_7 = \frac{1}{12}$, $\alpha_8 = \frac{1}{24}$, $\alpha_{17} = \frac{1}{120}$,} & {\small $\beta_3 = 0$, $\beta_4 = 0$,} \nline
        & & {\small $\frac{1}{4}\alpha_9 - \alpha_{10} + \alpha_{13} = 0$,} & {\small $\beta_5 = -\frac{1}{20} + \frac{1}{4}\alpha_9$,} \nline
        & & {\small $\frac{1}{20} + \frac{1}{2}\alpha_9 - 3\alpha_{10} + 3\alpha_{11} - \alpha_{14} = 0$,} & {\small $\beta_6 = -\frac{1}{40} - \frac{1}{8}\alpha_9 + \frac{1}{2}\alpha_{10}$,} \nline
        & & {\small $\frac{1}{40} + \frac{1}{2}\alpha_9 - 2\alpha_{10} + \alpha_{11} + \alpha_{12} - \alpha_{15} = 0$,} & {\small $\beta_7 = -\frac{1}{60} + \frac{1}{4}\alpha_9 - \alpha_{10} + \alpha_{11}$,} \nline
        & & {\small $\frac{1}{60} - \alpha_{11} + 2\alpha_{12} - \alpha_{16} = 0$.} & {\small $\beta_8 = -\frac{1}{120} + \frac{1}{8}\alpha_9 - \frac{1}{2}\alpha_{10} + \alpha_{12}$.} \nline
    \end{tabular}
    \caption{Effective order $q$, classical order $p$ conditions on $ \alpha $ and $ \beta $ for the main and starting methods, $M$ and $S$ respectively.}
    \label{tab:effective_OCs}
\end{table}


% \begin{example}\label{ex:Effective_RK32}
% Constructing an $s$-stage method of effective order 3 with classical order 2 requires satisfying the three conditions given in Table~\ref{tab:Effective_oc}. This is one less condition than constructing a method with classical order 3, since $\alpha_3$ is here a free parameter. After choosing a method and thus a value for $\alpha_3$, the order conditions for the starting method are now fixed as well.
% \end{example}
