\documentclass[10pt,a4paper,oneside]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	Preamble	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% packages %%%
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{enumerate,cite}
%\usepackage[toc,page]{appendix}
%\usepackage{diagrams}
%\usepackage[pdftex]{graphicx}
\usepackage{multirow}
\usepackage{url}
%\usepackage{nomencl}
\usepackage[usenames,dvipsnames]{color}
\usepackage{array}
\usepackage{arydshln}
\usepackage[dvips]{graphicx}
\usepackage{pstricks,pst-node,pst-tree}
%\usepackage{cm-super}
%\usepackage{ifthen}
\usepackage{subfig}
\usepackage{caption}
\usepackage{verbatim}
%\usepackage{maplestd2e}
\usepackage{slashbox}
%\usepackage[numbered]{mcode}
%\usepackage{rotating}
\usepackage{booktabs}
%\usepackage[font=small,format=plain,labelfont=bf,up,textfont=it,up]{caption}
\usepackage{epsfig}
\usepackage{bm}
%\usepackage{calc}
%\usepackage{pbox}
\usepackage{multicol}
\usepackage{xkeyval}
\usepackage[textsize=footnotesize]{todonotes}
\definecolor{darkred}{rgb}{.6,0,0}
\definecolor{darkblue}{rgb}{0,0,0.5}
\usepackage[colorlinks=true,citecolor=darkblue,urlcolor=darkblue,linkcolor=darkred]{hyperref}


%%% Page layout %%%
\setlength{\textwidth}{157mm}   
\setlength{\textheight}{225mm}
\setlength{\oddsidemargin}{-15mm} 
\setlength{\evensidemargin}{-2mm}
\setlength{\topmargin}{0mm}
\setlength{\marginparwidth}{35mm}

%%% Page style %%%
%\pagestyle{plain}

%%% New environments %%%
\newenvironment{biggertrees}{%
    \psset{treesep=3ex,levelsep=3ex,treemode=U,showbbox=false}
    \renewcommand\psedge{\ncline[linewidth=1.0pt]}
    \renewcommand{\Ts}{\Tdot[dotstyle=*,dotscale=1.1]}
}{}

%%% New commands %%%
\newcommand{\pd}[2]{\frac{\partial{#1}}{\partial{#2}}}
\newcommand{\pdseconda}[2]{\frac{\partial^2{#1}}{\partial{{#2}^2}}}
\newcommand{\pdsecondb}[3]{\frac{\partial^2{#1}}{\partial{#2}\partial{#3}}}
\newcommand{\Dt}{\Delta t}
\newcommand{\rowscell}[2]{\parbox{21pt}{\centering #1 \\ #2}}
\newcommand{\treecell}[2]{\parbox{26pt}{\centering #1 \\\vspace{3pt} #2}}
\newcommand{\nline}{\tabularnewline}
\newcommand{\colintodo}[1]{\todo[color=orange!20]{\textbf{C:} #1}}
\newcommand{\davidtodo}[1]{\todo[color=cyan!20]{\textbf{D:} #1}}
\newcommand{\yiannistodo}[1]{\todo[color=green!20]{\textbf{Y:} #1}}
\newcommand{\colincomment}[1]{\textcolor{blue}{\\\textbf{\footnotesize #1}\\}}
\newcommand{\davidcomment}[1]{\textcolor{red}{\\\textbf{\footnotesize #1}\\}}
\newcommand{\yianniscomment}[1]{\textcolor{OliveGreen}{\\\textbf{\footnotesize #1}\\}}
\newcommand{\prevtext}[1]{\begin{minipage}{140mm}
															{\footnotesize\newline\underline{Previous text}:\\#1}
												 \end{minipage}}
%\newcommand{\todo}[1]{\textbf{\red #1}}
\newcommand{\sspcoef}{\mathcal{C}}
\newcommand{\ceff}{\sspcoef_{\text{eff}}}
\newcommand{\clin}{\sspcoef_{\text{lin}}}
\renewcommand{\newline}{\vspace{10pt}}
\renewcommand{\comment}[1]{\textcolor{gray}{\\#1\\}}
\renewcommand{\indent}{\hspace{15pt}}

%%% New theorems %%%
\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{lemma}[theorem]{Lemma}
%\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{result}[theorem]{Result}
\newtheorem{remark}{Remark}

%%% Miscellaneous %%%
%\setcounter{secnumdepth}{3}
%\setcounter{tocdepth}{3}
%\setcounter{table}{0}
%\numberwithin{theorem}{section}
%\numberwithin{equation}{section}
%\numberwithin{definition}{section}
\numberwithin{theorem}{section}
%\numberwithin{example}{section}
\numberwithin{table}{section}
\numberwithin{figure}{section}
\numberwithin{equation}{section}
\newcolumntype{M}[1]{>{\vspace{2pt}\centering\hspace{0pt}\vspace{2pt}}m{#1}}
\definecolor{lightgray}{gray}{0.5}
%\makenomenclature
%\baselineskip=18pt plus1pt




\title{Effective order strong stability preserving Runge--Kutta methods}

\author{%
  Yiannis Hadjimichael\footnote{ 4700 King Abdullah University of
    Science \& Technology, Thuwal 23955, Saudi Arabia
    (\url{yiannis.hadjimichael@kaust.edu.sa},
    \url{david.ketcheson@kaust.edu.sa}).  The work of these authors
    was supported by funding from King Abdullah University of Science
    and Technology (KAUST).  } \and
  %
  Colin B.~Macdonald\thanks{Mathematical Institute, University of
    Oxford, OX1\,3LB, UK (\url{macdonald@maths.ox.ac.uk}).  The work
    of this author was supported by NSERC Canada and by Award No
    KUK-C1-013-04 made by King Abdullah University of Science and
    Technology (KAUST).} \and
  %
  David I.~Ketcheson\footnotemark[1] \and
  %
  J.~H.~Verner\thanks{Department of Mathematics, Simon Fraser
    University, Burnaby, British Columbia, V5A\,1S6 Canada
    (\url{jverner@pims.math.ca}).  The work of this author was
    supported by a grant from NSERC Canada.}  }



%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%		Actual document		%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\maketitle

\input{trees} % creates trees

\begin{abstract}
  In this work we apply the algebraic theory of Runge--Kutta methods
  and specifically the concept of effective order to strong stability
  preserving Runge--Kutta methods.  These methods of effective order
  satisfy relaxed order conditions when combined with special starting
  and stopping methods.  We construct such methods up to effective
  order four and establish their optimality (many methods achieve a
  previously theoretical upper bound on the SSP coefficient).  In
  particular, we present four-stage effective order four SSP RK
  methods which is not possible with four stages using classical
  order.  Numerical experiments demonstrate the feasibility of
  effective order in practice.  We prove that methods of effective
  order five cannot have strictly positive weights $b$.  A direct
  corollary is the non-existence of effective order five SSP methods.
\end{abstract}


\input{1_Intro}
\input{2_SSP}
\input{3_Algebraic_RK_new}
\input{4_ExRK_barrier}
\input{5_Optimal_ESSPRK}
\input{6_Numerics}
\input{7_Conclusion}


%\section*{Acknowledgments}
%The second author thanks Jim Verner for introducing him to this
%problem and for his support over the years.


%\appendix
%\include{AppendixA}

\bibliographystyle{acm} %apalike
\bibliography{bibliography}

\end{document}
