\documentclass[10pt,a4paper,oneside]{article}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Preamble %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%	

%%% packages %%%
\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage[usenames,dvipsnames]{color}
\usepackage[dvips]{graphicx} % [pdftex]
\usepackage{pstricks,pst-node,pst-tree}
\usepackage{bm}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{enumerate,cite}
\usepackage{url}
\usepackage{array}
\usepackage{arydshln}
\usepackage{subfig}
\usepackage{caption} % [font=small,format=plain,labelfont=bf,up,textfont=it,up]
\usepackage{verbatim}
\usepackage{rotating}
\usepackage{slashbox}
\usepackage{booktabs}
\usepackage{xkeyval}
\usepackage[textsize=footnotesize]{todonotes}
\definecolor{lightgray}{gray}{0.5}
\definecolor{darkred}{rgb}{.6,0,0}
\definecolor{darkblue}{rgb}{0,0,0.5}
\usepackage[colorlinks=true,citecolor=darkblue,urlcolor=darkblue,linkcolor=darkred]{hyperref}
\usepackage[normalem]{ulem}
\usepackage{accents}
%\usepackage[toc,page]{appendix}
%\usepackage{diagrams}
%\usepackage{nomencl}
%\usepackage{cm-super}
%\usepackage{ifthen}
%\usepackage{maplestd2e}
%\usepackage[numbered]{mcode}1
%\usepackage{epsfig}
%\usepackage{calc}
%\usepackage{pbox}

%%% Page layout %%%
\setlength{\textwidth}{160mm}   
\setlength{\textheight}{220mm}
\setlength{\oddsidemargin}{0mm} 
\setlength{\evensidemargin}{-5mm}
\setlength{\topmargin}{0mm}
\setlength{\marginparwidth}{35mm}
\setlength{\parindent}{5mm}
\setlength{\footskip}{20mm}     
\setlength\footnotesep{4mm}


%%% Page style %%%
\pagestyle{plain}

%%% New commands %%%
\newcommand{\sspcoef}{\mathcal{C}}
\newcommand{\ceff}{\sspcoef_{\textnormal{eff}}}
%\newcommand{\clin}{\sspcoef_{\textnormal{lin}}}
\newcommand{\clin}{\sspcoef^{\textnormal{lin}}_{s,q}}
%\newcommand{\clin}{R_{s,q}}
\newcommand{\pd}[2]{\frac{\partial{#1}}{\partial{#2}}}
\newcommand{\pdseconda}[2]{\frac{\partial^2{#1}}{\partial{{#2}^2}}}
\newcommand{\pdsecondb}[3]{\frac{\partial^2{#1}}{\partial{#2}\partial{#3}}}
\newcommand{\Dt}{\Delta t}
\newcommand{\DtFE}{\Delta t_{\textnormal{FE}}}
\newcommand{\equival}[1]{\mbox{{\large $\underaccent{\hspace{8pt}#1}{\simeq } \,$}}}
\newcommand{\rowscell}[2]{\parbox{21pt}{\centering #1 \\ #2}}
\newcommand{\treecell}[2]{\parbox{26pt}{\centering #1 \\\vspace{3pt} #2}}
\newcommand{\nline}{\tabularnewline}
\newcommand{\colintodo}[1]{\todo[color=orange!20]{\textbf{C:} #1}}
\newcommand{\davidtodo}[1]{\todo[color=cyan!20]{\textbf{D:} #1}}
\newcommand{\yiannistodo}[1]{\todo[color=green!20]{\textbf{Y:} #1}}
\newcommand{\colincomment}[1]{\textcolor{blue}{\\\textbf{\footnotesize #1}\\}}
\newcommand{\davidcomment}[1]{\textcolor{red}{\\\textbf{\footnotesize #1}\\}}
\newcommand{\yianniscomment}[1]{\textcolor{OliveGreen}{\\\textbf{\footnotesize #1}\\}}

%%% Renew commands %%%
\renewcommand{\newline}{\vspace{10pt}}
\renewcommand{\comment}[1]{\textcolor{gray}{\\#1\\}}
\renewcommand{\indent}{\hspace{15pt}}

%%% New theorems %%%
\newtheorem{theorem}{Theorem}[section]
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{result}[theorem]{Result}
\newtheorem{remark}[theorem]{Remark}

%%% Miscellaneous %%%
\numberwithin{theorem}{section}
\numberwithin{equation}{section}
\numberwithin{table}{section}
\numberwithin{figure}{section}
\newcolumntype{M}[1]{>{\vspace{2pt}\centering\hspace{0pt}\vspace{2pt}}m{#1}}
%\numberwithin{theorem}{section}
%\numberwithin{equation}{section}
%\numberwithin{definition}{section}
%\numberwithin{example}{section}
%\setcounter{secnumdepth}{3}
%\setcounter{tocdepth}{3}
%\setcounter{table}{0}
%\baselineskip=18pt plus1pt
%\makenomenclature

%%% General info %%%
\title{Explicit strong stability preserving Runge--Kutta methods with effective order of accuracy}
\author{
	Yiannis Hadjimichael\footnote{4700 King Abdullah University of Science \& Technology, 
	(KAUST), Mathematical and Computer Sciences and Engineering Division, Thuwal 23955, 
	Saudi Arabia
	(\url{yiannis.hadjimichael@kaust.edu.sa}, 
	\url{david.ketcheson@kaust.edu.sa}).
	The work of these authors is supported by Award No. FIC/2010/05, made by King 
	Abdullah University of Science and Technology (KAUST).}
    	\and 
    	Colin B.~Macdonald\thanks{Mathematical Institute, University of Oxford, OX1\,3LB, UK 
    	(\url{macdonald@maths.ox.ac.uk}).
    	The work of this author was supported by NSERC 
    	Canada and by Award No KUK-C1-013-04 made by King Abdullah University of Science 
    	and Technology (KAUST).}
    	\and 
    	David I.~Ketcheson\footnotemark[1]
    	\and 
    	Jim.~H.~Verner\thanks{Department of Mathematics, Simon Fraser University,
    	Burnaby, British Columbia, V5A\,1S6, Canada
    	(\url{jverner@pims.math.ca}).
    	The work of this author was supported by a grant from NSERC Canada.}
}


%%%%%%%%%%%%%%%%%%%%%%%%%%%% Main document %%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

	\maketitle
	
	%%% Supporting macros %%%
	\input{trees} % creates trees
	
	\begin{abstract}
  		We apply the concept of effective order to strong stability preserving 
  		(SSP) explicit Runge--Kutta methods.
  		Relative to classical Runge--Kutta methods, methods of effective order 
  		are designed to satisfy a relaxed set of order conditions, but yield higher 
  		order accuracy  when composed with special starting and stopping methods. 
         The relaxed order conditions allow for greater freedom in the design
         of methods with effective order of accuracy. 
         We show that this allows the construction of four-stage SSP methods with 
         effective order four (such methods cannot have classical order four). 
         However, we also prove that effective order five methods---like classical
         order five methods---require the use of non-positive weights and so cannot
         be SSP.
         By numerical optimization, we construct explicit SSP Runge--Kutta methods 
         up to effective order four and establish the optimality of many of them.
  		Numerical experiments demonstrate the validity of these methods in
  		practice.
	\end{abstract}

	%%% Main body %%%
	\input{1_Intro}
	\input{2_SSP}
	\input{3_Algebraic_RK}
	\input{4_ExRK_barrier}
	\input{5_Optimal_ESSPRK}
	\input{6_Numerics}
	\input{7_Conclusion}

	%\section*{Acknowledgments}{
	%The second author thanks Jim Verner for introducing him to this
	%problem and for his support over the years.
	%}

	%%% Appendix and Bibliography %%%
	%\appendix
	%\include{AppendixA}
	\bibliographystyle{acm} %apalike
	\bibliography{bibliography2}

\end{document}
